Binning Strategies and Related Loss for Large Data

### 2015-02-18 Paper.Rnw
quantile binning edits

\begin{figure}[hbtp]
  \centering
  \subfloat[Standard Binning]{\includegraphics[keepaspectratio=true, width=3in]{./figure/Binning-DataAndBin.pdf}}
%   \subfloat{\includegraphics[keepaspectratio=TRUE, width=2in]
%   {./figure/RandomBinningDataPoints1}}
  \subfloat[Random Binning]{\includegraphics[keepaspectratio=TRUE, width=3in]
  {./figure/RandomBinningDataPoints2.pdf}}
  \caption{Visualization of how data pairs are assigned to bins centered at (5, 25) and (15,25) under the two rectangular binning algorithms. Points in the standard algorithm are contained within the defined bin region, where points from the random algorithm may be assigned to bins from outside the defined bin regions.}
  \label{fig:BinningTutorial}
\end{figure}

@@

Suppose that we have values $c_j$ for $j \in \{1,2,...J\}$ that we wish to divide into quantile based bins. The number of quantile based bins, $N_q$ , must be defined in order to assign values to bin centers. The quantile binning function $b^q_C(\cdot)$ assigns ordered values $c_{(j)}$ to bin centers $c^\ast_{(j)}$ as 
%
\begin{eqnarray}\label{quantbin}
c^\ast_{(j)} = b^q_C(c_{(j)}) = c_{(\eta)} \text{, where } \eta = \left\lfloor j/ \left\lceil J/N_c \right\rceil \right\rfloor \left\lceil J/N_c \right\rceil ,
\end{eqnarray}
%
\noindent and $\left\lceil \cdot \right\rceil$ is the ceiling function that rounds any real number up to the next highest integer.  To better understand this function we notice that $\left\lceil J/N_c \right\rceil$ represents the number of values to be assigned to each quantile bin, $n_q$, rounded up; thus $\eta$ is the index for the ordered value where the next lower quantile based bin will be centered.

>>>>

\begin{figure}[h]
  \centering
  \subfloat[Standard Binning]{\includegraphics[keepaspectratio=true, width=3in]{./figure/Binning-DataAndBin.pdf}}
%   \subfloat{\includegraphics[keepaspectratio=TRUE, width=2in]
%   {./figure/RandomBinningDataPoints1}}
  \subfloat[Random Binning]{\includegraphics[keepaspectratio=TRUE, width=3in]
  {./figure/RandomBinningDataPoints2.pdf}}
  \caption{Visualization of how data pairs are assigned to bins centered at (5, 25) and (15,25) under the two rectangular binning algorithms. Points in the standard algorithm are contained within the defined bin region, where points from the random algorithm may be assigned to bins from outside the defined bin regions.}
  \label{fig:BinningTutorial}
\end{figure}

@@

Suppose that we have values $c_j$ for $j \in \{1,2,...J\}$ that we wish to divide into quantile based bins. Defining the number of quantile based bins, $K$, results in a set of possible quantile bin centers. It follows that the bin boundaries are $Q_{k/K}$ for $k \in \{1,\dots,K\}$, where $Q_p$ is the the $p^{th}$ empirical quantile of value using the inverse empirical distribution function. \ktm{Note this is the quantile function as defined in R; find related literature source.} The bin centers are defined as the midpoint of each consecutive pair of bin boundaries $\{c^\ast_k = (Q_{k-1/K} + Q_{k/K})/2  \text{ for } k=0,\dots,K \}$. The quantile binning function $b^q_C(\cdot) : (c_j) \rightarrow c^\ast_k$ is defined as 
%
\begin{eqnarray}\label{quantbin}
b^q_C(c_j) = (Q_{k-1/K} + Q_{k/K})/2 \text{ where } Q_{k-1/K} < c_j \le Q_{k/K}.
\end{eqnarray}
%
The two dimensional analog for quantile binning is not explored, as this binning strategy will be used in stictly the univariate setting for this manuscript.
